---
title: "Getting Started with hierviz"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with hierviz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hierviz)
```

## Introduction

`hierviz` is designed for working with hierarchical data structures, particularly when you need to:

1.  Flatten nested categories into a more analyzable format
2.  Visualize trends across categories
3.  Analyze changes between periods

## Working with Hierarchical Data

### Flattening Categories

The `flatten()` function transforms hierarchical categories into full names. This is useful when your data has sections and items:

```{r}
# Create sample hierarchical data
hier_data <- tibble::tibble(
  Category = c(
    "Hardware:", 
      "Surface",
      "Xbox",
      "Peripherals",
    "Software:",
      "Office",
      "Windows",
      "Azure"
  ),
  Value = c(
    NA,    # Hardware header
      100,
      200,
      150,
    NA,    # Software header
      300,
      400,
      500
  )
)

# Flatten the categories
flat_data <- flatten(hier_data)
print(flat_data)
```

Header rows (ending in ":") are removed. Categories are combined with their section headers, but ther structure is preserved in the full category names.

### Creating Queries

The `query()` function is an abstraction on what you want to visualize or analyze:

```{r}
# Create sample data with multiple periods
time_data <- tibble::tibble(
  Category = c("A", "B", "C"),
  Q1 = c(100, 200, 300),
  Q2 = c(110, 180, 320),
  Q3 = c(120, 190, 310),
  Q4 = c(130, 200, 330)
)

# Different ways to create queries
q1 <- query(time_data)                            # Use defaults
q2 <- query(time_data, "Category")                # Specify category column
q3 <- query(time_data, "Category", c("Q1", "Q2")) # Specify both columns
```

## Visualization

### Line Charts

The `line_chart()` function visualizes trends across categories:

```{r}
# Create a line chart showing all periods
line_chart(q1)
# Compare specific periods
line_chart(q3)
```

Each category gets its own color. Lines connect values across periods. Points mark actual values.

### Change Tables

The `change_table()` function shows period-over-period changes:

```{r}
# Create a change table
change_table(q1)
```

Green indicates positive changes. Red indicates negative changes. Black indicates no changes. Changes are shown as percentages.

### Custom Comparisons

You can choose specific periods to compare:

```{r}
# Compare Q1 and Q4 only
q_custom <- query(time_data, "Category", c("Q1", "Q4"))
change_table(q_custom)
```

## Best Practices

1.  **Data Structure**
    -   Use consistent delimiters (":") for section headers
    -   Ensure numeric columns contain actual numbers
2.  **Category Names**
    -   Avoid using ":" in category names (reserved for headers)
    -   Keep names concise but descriptive
3.  **Visualization**
    -   Consider filtering to most important categories if you have many

## Common Issues

1.  **Non-Numeric Data**

    ``` r
    # This will error due to non-numeric data column
    bad_nums <- tibble::tibble(
      Category = c("A", "B"),
      Value = c("100", "text")  # Should be numeric
    )
    query(bad_nums)  # Error: Column must contain numeric values
    ```
